---
title: "phecopack: exploring PheCode for ICD10"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{phecopack: exploring PheCode for ICD10}
  %\VignetteEncoding{UTF-8}
output:
  rmarkdown::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

From [a Journal of Medical Internet Research article
on PheCode mapping](https://medinform.jmir.org/2019/4/e14325/):

> The initial version of phecodes consisted of 733 custom groups of ICD Ninth Revision, Clinical Modification (ICD-9-CM) diagnosis codes. The most recent iteration of phecodes consists of 1866 hierarchical phenotype codes that map to 15,558 ICD-9-CM codes [13,14]. However, many health systems and international groups use the International Classification of Diseases, 10th Revision (ICD-10) or the International Classification of Diseases, 10th Revision, Clinical Modification (ICD-10-CM) codes [15], therefore necessitating a new phecode map.

The `phe.icd10` data frame in this package
was retrieved from the [PheWAS catalog site](https://phewascatalog.org/phecodes_icd10) on May 6 2024.
The version is listed as v1.2 beta.  

The publication noted above
is accessible in [PMC](https://pubmed.ncbi.nlm.nih.gov/31553307/).

# Basic resources

## The basic table

This table can be searched interactively.

```{r lk1}
library(phecopack)
data(phe.icd10)
DT::datatable(phe.icd10[,c("ICD10", "ICD10.String",
  "Phenotype", "PheCode")])
```

## The search function

```{r lk2}
chk = get_icd10(c("asthma*", "cardiac", "malignant"))
lapply(chk, function(x) head(x[, c("ICD10.String", "Phenotype")]))
```
